Topic modeling on EU and the USA research projects

### 2017-06-20 thesis_report/draft_thesis.tex
Fixed: misspellings in the Related work section

D.Blei's research focuses on probabilistic topic models, Bayesian methods and
approximate posterior inference. In his work "Introduction to Probabilistic
Topic Models",\cite{Blei11introductionto} he describes the basic ideas behind
LDA and applies the model to articles from the journal "Science". The model
assumes that the topics are generated first, before the documents and that the
documents axhibit multiple topics, but each one of them exhibit those topics in
different proportion. In our implementation we also use the LDA model and we
follow a similar approach, since we don't use any previous information about the
subjects that run through our dataset and the documents are not labeled with
keywords or topics. The topic distributions arise by computing the hidden
structure that resembles the thematic structure of the collection.

The paper that mostly infuenced our work is "Finding Scientific Topics", by
T.Griffiths and M.Steyvers.\cite{griffiths_steyvers04} They investigated
abstract data from articles published in the Proceedings of the National Academy
of Sciences (PNAS) from 1991 to 2001 and compared research topics obtained from
topic modeling with existing categories. We also use the abstacts of the
projects as documents for the model. (to be continued)

Griffiths and Steyvers describe LDA and present a Marckv Chain Monte Carlo
algorithm for inference in the model, illustrating the operation of the
suggested algorithm on a small dataset, and then on a larger corpus of PNAS
abstracts. Finally, they define the dynamics of the topics as a means of gaining
insight into the dynamics of science. As "hot" topics are identified the topics
that appear more frequently over the years (positive linear trend), and as
"cold" the ones that appear less frequently (negative linear trends over time).
We also analyze the topics over time, in order to highlight potential trends in
scientific research in EU and USA.

>>>>

D.Blei's research focuses on probabilistic topic models, Bayesian methods and
approximate posterior inference. In his work "Introduction to Probabilistic
Topic Models",\cite{Blei11introductionto} he describes the basic ideas behind
LDA and applies the model to articles from the journal "Science". The model
assumes that the topics are generated first, before the documents and that the
documents exhibit multiple topics, but each one of them exhibit those topics in
different proportion. In our implementation we also use the LDA model and we
follow a similar approach, since we don't use any previous information about the
subjects that run through our dataset and the documents are not labeled with
keywords or topics. The topic distributions arise by computing the hidden
structure that resembles the thematic structure of the collection.

The paper that mostly infuenced our work is "Finding Scientific Topics", by
T.Griffiths and M.Steyvers.\cite{griffiths_steyvers04} They investigated
abstract data from articles published in the Proceedings of the National Academy
of Sciences (PNAS) from 1991 to 2001 and compared research topics obtained from
topic modeling with existing categories. In our implementation, we also use 
the abstacts of the projects as documents for the model.

Griffiths and Steyvers describe LDA and present a Markov Chain Monte Carlo
algorithm for inference in the model, illustrating the operation of the
suggested algorithm on a small dataset, and then on a larger corpus of PNAS
abstracts. Finally, they define the dynamics of the topics as a means of gaining
insight into the dynamics of science. As "hot" topics are identified the topics
that appear more frequently over the years (positive linear trend), and as
"cold" the ones that appear less frequently (negative linear trends over time).
We also analyze the topics over time, in order to highlight potential trends in
scientific research in EU and USA.
